On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-08-17 Manuscript/manuscript.tex
minor edits

\subsection{2022 World Championships}\label{sec:dataworld}

@@

For analysis purposes, we will often pool reaction times of men's final and 
men's semifinal heats together and refer to this data as the ``pooled data"
throughout the paper. 
%\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
%use it, I do reference it in the methods section}
The pooled data increases the sample size which is advantageous because in some
years there are very few final heat observations.  For example, in 2022 there
were only five data points due to two disqualifications and one athlete who 
did not compete.  We also look at data that does not include 2022 with year values
between 1999 and 2019.  Thus, we have four datasets with their respective naming
in parentheses: the final's data from 1999 to 2019 (Old Finals), final's data 
from 2019 to 2022 (All Finals), pooled data from 1999 to 2019 (Old Pooled), and 
pooled data from 1999 to 2022 (All Pooled).

The data under consideration is best visualized using side-by-side boxplots, as
shown in Figure~\ref{fig:Boxplots}.  It is clear that the reaction times in the 
2022 boxplot for both 
the men's finals and men's pooled semifinal and final data are lower (indicating 
faster reaction times)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
\citet{brosnan2017effects} found median reaction times of 0.156 seconds 
for men when looking at data from 1999 to 2014.

We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
differ from year to year. The venue of the World Championships changes each 
year, so weather or climate related factors such as humidity, precipitation, 
elevation, may indeed play a role, but additionally, both the technology and 
false start penalization rules have changed over the study period.

@@

\subsection{Beyond 2022 World Championships}\label{sec:databeyond}
The motivation for looking beyond World Championships data came from an 
exploratory analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.

@@

\subsection{Rank-based Comparison}\label{sec:rank_methods}

@@

A summary of the two effect model is shown in table~\ref{tab:Gamma_parameters}
for the pooled datasets and shows the standard error of the year
effect increases for both the men's final and men's pooled data sets when
including 2022. We choose to only look at the pooled datasets, as R displayed
warnings when creating the finals only models.  The two effect model with venue
and heat effect was the best model, as proven by an ANOVA test and decrease in
AIC.  For the entire pooled dataset; the AIC for the venue effect only model was
$-1847.9$, the heat effect only model AIC was $-1894.4$ and the both effect model,
which we will present throughout this section AIC was $-1898.9$. Thus, the two
effect model is superior and will be presented throughout the rest of the paper.
\of{How specific do I need to be in this section? Will people who read this paper
run the code to see the p value or should I say it directly and how do I say it?}

@@

\jy{use two-sided}
\of{Resolved}

\jy{include sample sizes; permutation number (B);}
\of{Resolved}

>>>>

\subsection{World Championships 1999--2022}\label{sec:dataworld}

@@

For analysis purposes, we will often pool reaction times of men's final and 
men's semifinal heats together and refer to this data as the ``pooled data"
throughout the paper. 
%\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
%use it, I do reference it in the methods section}
The pooled data increases the sample size which is advantageous because in some
years there are very few final heat observations.  For example, in 2022 there
were only five data points due to two disqualifications and one athlete who 
did not compete.  We also look at data that does not include 2022 with year values
between 1999 and 2019.  Thus, we have four datasets with their respective naming
\jy{Clean up}
in parentheses: the final's data from 1999 to 2019 (Old Finals), final's data 
from 2019 to 2022 (All Finals), pooled data from 1999 to 2019 (Old Pooled), and 
pooled data from 1999 to 2022 (All Pooled).

The data under consideration is best visualized using side-by-side boxplots, as
shown in Figure~\ref{fig:Boxplots}.  It is clear that the reaction times in the 
2022 boxplot for both 
the men's finals and men's pooled semifinal and final data are lower (indicating 
faster reaction times)
than in previous years.  The median for the men's pooled data in 2022 was 0.129
and the median for the men's final data in 2022 was 0.126.  For comparison,
{brosnan2017effects} found median reaction times of 0.156 seconds 
for men when looking at data from 1999 to 2014.
We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
differ from year to year. The venue of the World Championships changes each 
year, so weather or climate related factors such as humidity, precipitation, 
elevation, may indeed play a role, but additionally, both the technology and 
false start penalization rules have changed over the study period.

@@

\subsection{Comparison Data with 2022 World Championships}\label{sec:databeyond}
The motivation for looking beyond World Championships data came from an 
exploratory analysis examining how United States athletes reacted at the 
2022 United States Track and Field Championships.

@@

\subsection{Rank-based Comparison}\label{sec:rank}

@@

A summary of the two effect model is shown in Table~\ref{tab:Gamma_parameters}
for the pooled datasets and shows the standard error of the year
effect increases for both the men's final and men's pooled data sets when
including 2022. We choose to only look at the pooled datasets, as R displayed
warnings when creating the finals only models.  The two effect model with venue
and heat effect was the best model, as proven by an ANOVA test and decrease in
AIC.  For the entire pooled dataset; the AIC for the venue effect only model was
$-1847.9$, the heat effect only model AIC was $-1894.4$ and the both effect model,
which we will present throughout this section AIC was $-1898.9$. Thus, the two
effect model is superior and will be presented throughout the rest of the paper.
\of{How specific do I need to be in this section? Will people who read this paper
run the code to see the p value or should I say it directly and how do I say it?}

@@
